---
title: "Homework 1"
author: "Ethan T. Richardson"
date: "2023-01-19"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(mdsr)
```

This $#!? is automated


### Question 1

It is done, I'm not sure I can show it in a code section here but I'll show you a mathematical P

$P$

Pretty cool, right?

The above is before you explained how to prove it, but I laughed enough to keep it

Here's the URL: https://github.com/erichardson2/DSCI101.git

### Question 2

```{r}
cubs <- read.csv("https://raw.githubusercontent.com/gjm112/DSCI101/main/data/cubs_all_time.csv")
```

#### Part A

```{r}
#Sum of the amount of games the Cubs lost
won <- sum(cubs$W)

#Sum of the amount of games the Cubs lost
lost <- sum(cubs$L)
```

The Cubs won a total of `r won` games and lost a total of `r lost` games between 1876 and 2022.

#### Part B

```{r}
#Season with most runs
mostYear <- cubs$Year[which(cubs$R == max(cubs$R))]

#Season with least runs
leastYear <- cubs$Year[which(cubs$R == min(cubs$R))]
```

The Cubs scored the most runs in `r mostYear` and the least runs in `r leastYear`. 

2020 was a very depressing season.

#### Part C

```{r}
#filter dataframe by years that attendance over 3mill
attHigh <- filter(cubs, Attendance > 3000000)

amount <- length(attHigh$Attendance)
```
The Cubs had `r amount` seasons that were above 3 million in total attendance


#### Part D

```{r}
cubsRatio <- mutate(cubs, ratio = R/RA)

maxRatio <- max(cubsRatio$ratio)
```

The greatest ratio of runs for to runs against between 1982 and 2000 was about `r maxRatio`:1

## Question 3

```{r}
college <- read.csv("https://raw.githubusercontent.com/gjm112/DSCI101/main/data/top_colleges_2022.csv")
```

#### Part A

```{r}
collegeIll <- filter(college, state == 'IL')

amountIll <- length(collegeIll$state)
```

There are `r amountIll` colleges from this data set that are in Illinois

#### Part B

```{r}
underPopLoy <- college$undergradPop[which(college$organizationName == 'Loyola University Chicago')]

higher <- college %>%
  filter(undergradPop > underPopLoy) %>%
  summarize(length(undergradPop))
```

There are `r higher` colleges with a larger undergraduate student population than Loyola University Chicago


#### Part C

```{r}
#Latitude is North and South
North <- college %>%
  summarize(organizationName,latitude, NA.rm = TRUE) %>%
  arrange(desc(latitude))

South <- college %>%
  summarize(organizationName,latitude, NA.rm = TRUE) %>%
  arrange(latitude)

```
The college furthest North is `r North[1,1]`. While Alaska is the state furthest to the north, they are not represented in this data set. As such, Washington has the majority of the most northern colleges

The college furthest South is `r South[1,1]`. Hawaii is the only official US state below the Tropic of Cancer


```{r}
#Longitude is East and West
East <- college %>%
  summarize(organizationName,longitude, NA.rm = TRUE) %>%
  arrange(desc(longitude))

West <- college %>%
  summarize(organizationName,longitude, NA.rm = TRUE) %>%
  arrange(longitude)
```

While the most Eastern University according strictly to the original data set says it's `r East[1,1]` due to a typo, it is actually `r East[2,1]`.

The most Western college is `r West[1,1]`, which makes the college also the most South-Western college in the data set, what a fun fact.
